---
title: "STA 426 - Week 1"
author: "David Wissel"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: bookdown::html_document2

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

# Setup
```{r bookkeeping, message=FALSE}
library(tidyverse)
theme_set(theme_minimal())
mu <- 1
sigma <- 0.25
n <- 1e2
seed <- 42
decimal_places <- 3
# Found through trial and error to make histograms look presentable
bin_width <- 0.18
```

We set the parameters explicitly such that we can easily modify them and change them later if a different distribution was desired. We then simulate the data:

```{r sim}
set.seed(seed)
dat <- data.frame(x = rlnorm(n, mu, sigma))
```


# Histograms
We shall use `ggplot2` to display histograms of both the original data and the log-transformed data.

## Non-transformed data


```{r hist, fig.cap = paste("Histogram of", n, "randomly generated observations from a log normal distribution with mean =", mu, "and sigma =", sigma)}
p <- dat %>%
  ggplot(aes(x)) +
  geom_histogram(binwidth = bin_width)

p
```


We can see in Figure \@ref(fig:hist) that the histogram of the untransformed data takes the skewed shape of a log-normal distribution.

## Log-transformed histogram


```{r histtrans, fig.cap = paste("Histogram of the log transformation of", n, "randomly generated observations from a log normal distribution with mean =", mu, "and sigma =", sigma)}
p + scale_x_log10(name = "log(x)")
```


As expected, Figure \@ref(fig:histtrans) shows that after log-transforming the data, the histogram takes (roughly) the shape of a normal distribution with mean `r mu` and standard deviation `r sigma`.

# Calculations
By using the functions `mean` and `var`, we can calculate that the mean and variance of the sample are `r round(mean(dat$x), decimal_places)` and `r round(var(dat$x), decimal_places)` respectively (each rounded to `r decimal_places` decimal places).
